Do More Subdivisions in Wards Increase Voter Turnout? An Analysis of the 2022 Toronto Municipal Election

### 2023-06-03 outputs/paper/paper.qmd
Adding more to the discussion section

## Discussion
Overall, the data reveals that having more subdivisions within wards does not automatically correlate to higher voter turnout, as illustrated by @fig-comparing-sub-turnout. Ward 4 (Parkdale-High Park) which had the highest voter turnout at 38% has 70 subdivisions within the ward, which falls approximately in the middle of the number of subdivisions across all 25 wards. Ward 7 (Humber River-Black Creek) had a 22% voter turnout and 47 subdivisions, emphasizing that having fewer subdivisions potentially contributed to having the lowest voter turnout across the city. Ward 10 (Spadina-Fork York), which had the second lowest voter turnout at 24%, accounts for the highest number of subdivisions within a ward at 94. Previous research reveals that younger, more transient communities reside in neighbourhoods in the Spadina-Fork York riding and are less likely to engage with municipal politics, potentially contributing to the low voter turnout despite the high number of subdivisions [@Turnout22].

Furthermore, Ward 19 (Beaches-East York) and Ward 14 (Toronto Danforth) which both saw the second highest voter turnout at 36% had 57 and 52 subdivisions respectively, which is on the lower end of number of subdivisions.

>>>>

## Discussion
Overall, the data reveals that having more subdivisions within wards does not automatically correlate to higher voter turnout, as illustrated by @fig-comparing-sub-turnout. Ward 4 (Parkdale-High Park) which had the highest voter turnout at 38% has 70 subdivisions within the ward, which falls approximately in the middle of the number of subdivisions across all 25 wards. Ward 19 (Beaches-East York) and Ward 14 (Toronto Danforth) which both saw the second highest voter turnout at 36% had 57 and 52 subdivisions respectively, once again highlighting that more subdivisions does not necessarily correlate to higher voter turnout.

Futhermore, Ward 7 (Humber River-Black Creek) had a 22% voter turnout and 47 subdivisions, emphasizing that having fewer subdivisions potentially contributed to having the lowest voter turnout across the city. Ward 10 (Spadina-Fork York), which had the second lowest voter turnout at 24%, accounts for the highest number of subdivisions within a ward at 94. Previous research reveals that younger, more transient communities reside in the Spadina-Fork York riding and are less likely to engage with municipal politics, potentially contributing to the low voter turnout despite the high number of subdivisions [@Turnout22].

The Wards within Scarborough (Wards 20-25) consistently had some of the lowest voter turnouts and fewest number of subdivisions across all 25 wards. Ward 23 (Scarborough North) witnessed the third lowest voter turnout across the entire city at 25% and the fewest number of subdivisions at 38. Ward 20 (Scarborough Southwest) had the highest turnout of all Scarborough wards at 32% and had 61 subdivisions. Scarborough's challenges with accessible public transit are well recognized, potentially contributing to the lower voter turnout.

The creation of subdivision boundaries within wards lacks transparency and clarity. The City Clerk is responsible for drawing up all ward subdivisions. We don’t know what data and information they use to make these decisions and what types of bias may be introduced. Don’t know if there’s a public input process or if like creating ward boundaries, it is an individual who has little local knowledge of the very community impacted by the creation of boundaries

Toronto Elections [@ElxnAccessibility] clearly outlines their process for finding and establishing accessible polling locations within subdivisions but does not elaborate on the process of establishing subdivisions themselves.

There is a lack of clarity whether the people who voted at Advanced Polls or by Mail In Ballot were removed from the total number of electors by subdivision, since these individuals did not vote at their assigned subdivision. This could have an impact on the total number of voters on a subdivision and ward basis.

There are several limitations with conducting this analysis. Number of subdivisions and polling location placement alone cannot solely dictate voter behaviour, as there are many other socio-economic factors which inform why an elector does or does not cast a ballot.
